---
title: "Reproducible Research: Course Project 1"
author: "Jun Yitt, Cheah"
date: "March 22, 2017"
output: html_document
---

#Loading and preprocessing the data
```{r, echo = T}
suppressMessages(library(dplyr))
suppressWarnings(library(ggplot2))


setwd("C:/Users/User/Desktop")
df <- read.csv("activity.csv", stringsAsFactors = F)
```


#What is mean total number of steps taken per day?

###1. Calculate the total number of steps taken per day
```{r, echo = T}
stepsday <- df %>% group_by(date) %>% summarise(sum = sum(steps, na.rm = T))
stepsday
```

###2. Make a histogram of the total number of steps taken each day
```{r, echo = T}
stepsday.v <- as.numeric(as.data.frame(stepsday)[,2])
hist(stepsday.v, breaks = 30)
```

###3. Calculate and report the mean and median of the total number of steps taken per day
```{r, echo = T}
mean1 <- format(mean(stepsday.v) )
median1 <- format(median(stepsday.v))
mean1
median1
```

The mean and median of the total number of steps taken per day are `r mean1` and `r median1` respectively.


# What is the average daily activity pattern?

###1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo = T}
stepsavg <- df %>% group_by(interval) %>% summarise(avg = mean(steps, na.rm = T))
plot(x=stepsavg$interval, y = stepsavg$avg, type = "l")
```

###2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo = T}
maxint <- stepsavg$interval[stepsavg$avg == max(stepsavg$avg)]
maxint
```
The `r maxint`th 5-minute interval contains the maximum number of steps.


#Imputing missing values

###1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo = T}
sumna <- sum(is.na(df$steps))
sumna
```
The total number of missing values in the dataset is `r sumna`.

###2. Devise a strategy for filling in all of the missing values in the dataset.
First, create a duplicate dataset of the original dataset.   
Then, replace each missing value with the mean of the total number of steps taken in that respective 5-minute interval across the dataset.

```{r, echo = T}
df2 <- df

uniqueinterval.v <- unique(df2$interval)
stepsavg.df <- as.data.frame(stepsavg)

for(x in uniqueinterval.v){
      u1 <- is.na(df2$steps)
      u2 <- df2$interval %in% x
      df2[u1&u2, "steps"] <- stepsavg.df[stepsavg.df$interval == x, "avg"]
}
```


###3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
The new dataset is created.
```{r, echo = T}
head(df2)
```

###4. Make a histogram of the total number of steps taken each day. Calculate and report the mean and median total number of steps taken per day.
```{r, echo = T}
stepsday2 <- df2 %>% group_by(date) %>% summarise(sum = sum(steps, na.rm = T))
stepsday2.v <- as.numeric(as.data.frame(stepsday2)[,2])

hist(stepsday2.v, breaks = 30)

mean2 <- format(mean(stepsday2.v))
median2 <- format(median(stepsday2.v))
mean2
median2

```

The new mean and median of the total number of steps taken per day are `r mean2` and `r median2` respectively. The new mean and median is slightly higher than the mean and median computed earlier (`r mean1` and `r median1` respectively). 
Hence, the estimates might be higher using the new dataset.



# Are there differences in activity patterns between weekdays and weekends?

###1. Creating a factor variable to distinguish weekday and weekend
```{r, echo = T}
df3 <- df2 %>% mutate(day1 = weekdays(as.Date(date)), wd = weekdays(as.Date(date)) %in% c("Saturday", "Sunday") )  %>% mutate(factorwk = factor(wd, c(FALSE, TRUE), labels = c("weekday", "weekend")))

head(df3)
tail(df3)
```                                                     
                                                                                                    
###2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)     
```{r, echo = T}                                          
df4 <- df3 %>% group_by(factorwk, interval) %>% summarise(msteps = mean(steps))

ggplot(data = df4, mapping = aes(x = interval, y = msteps)) + geom_line() + facet_grid(factorwk~.) + ylab("Average number of steps taken")
```      


